Evaluating computational methods for modeling off-normal operation of gas centrifuge cascades

### 2018-05-22 paper.tex
working on the flow table
first set of review comments

\usepackage[acronym]{glossaries}
\newacronym{UOX}{UOX}{uranium oxide fuel}
\newacronym{MOX}{MOX}{mixed oxide fuel}
\newacronym{LWR}{LWR}{light water reactor}
\newacronym{FBR}{FBR}{fast breeder reactor}

@@

Fuel cycle simulations, as with any simulation process, do not produce results
without errors.  Those errors have different sources: data uncertainties (from
the simulator's reliance on previously estimated/measured metrics), modeling
uncertainties (from simplifications made by the simulator) and the
``operational`` uncertainty (from the operation parameter uncertainties of
different facilities).

Simulation errors are important in the context of treaty verification. 
Some scenarios require estimating historical fissile
material production based on records of facility operation and material
transfer.  If the measurements of facility operation and material transfer are
subject to uncertainty, the total production quantities will then also be
uncertain, creating an opportunity for undetected diversion.  This study seeks
to understand which measures of facility operation have the most impact on the
uncertainty of fissile material production across a complete fuel cycle.
To accomplish this, a Total Monte Carlo methodology will be applied to the Cyclus
fuel cycle simulator \cite{cyclus} to estimate the effect of individual facility uncertainties on the output metrics.

@@

To estimate the uncertainty of fuel cycle output metrics, a Total Monte
Carlo approach has been applied.  In the following, output metric uncertainty
corresponds to the standard deviation of the output metrics from 199
different simulations.  For each deployed facility, the
parameters are normally distributed with an artificially large standard
deviation of $10\%$.

To pursue this work, both the Cycamore\cite{cycamore} package\footnote{Low
fidelity facilities package for Cyclus} and the
CyCLASS\footnote{Reactor and fuel fabrication facility based on CLASS\cite{CLASS} models} were
updated\cite{mouginot_2018, cyclass} to evaluate the uncertainty of several
operational parameters in their associated facilities.  Table
\ref{tab:package_uncertainty} summarizes all uncertainty-tracking modifications implemented in the
facilities of Cycamore and CyCLASS.

\begin{table}[htb]
\centering
  \caption{Summary of facility modifications according to their source package.}
\begin{tabular}{cll}
\toprule

Package   & Facility   & Parameters                \\
\midrule
Cycamore & Separation & Separation efficiency     \\
         & Storage    & Residence Time            \\
\midrule
CyCLASS  & Reactor    & Cycle Length              \\
         &            & Power                     \\
         &            & Fuel Enrichment (PWR-UOX) \\

@@

The uncertainty of the different parameters in this study
are systematic for each new deployed facility: each time a new facility is
deployed (such as a reactor or a fuel fabrication facility) a new set of
parameters are computed.

\subsection{The Fuel Cycle Scenario}
This work aims to assess the impact of the operational uncertainties of a simple
transition, inspired from the EG23 fuel cycle
of the Nuclear Fuel Cycle Evaluation and Screening Report\cite{ES}. Pictured in Figure \ref{fig:cycle}, this is a
transition from a light water reactor fleet loaded with \gls{UOX} fuel to a
sodium fast reactor fleet loaded \gls{MOX} fuel, considering a $1\%/$y growth
of the generated power (Figure \ref{fig:power}).
\begin{figure}[ht] % replace 't' with 'b' to force it to be on the bottom
  \centering
  \includegraphics[scale=0.31]{cycle.png}
  \caption{Illustration of the material flow between the different facilities,
  with each operational parameter associated with an
  uncertainty labeled in green.}\label{fig:cycle}
\end{figure}

As illustrated in Figure \ref{fig:power}, the transition starts with a fleet
composed of only \glspl{LWR} loaded with enriched \gls{UOX} fuel.  The actual
transition starts around year 35, with the deployment of \glspl{FBR}. The \glspl{FBR} are loaded
with \gls{MOX}, which consists of plutonium blended with natural
uranium.

\begin{figure}[ht] % replace 't' with 'b' to force it to be on the bottom
    \centering
    \includegraphics[scale=0.18]{power.png}
    \caption{Electric power generated over time, with the full production,
        the \gls{LWR} contribution, and the \gls{FBR} contribution labeled in blue, orange, and green, respectively.}
    \label{fig:power}
\end{figure}

@@

Both fuel reprocessing and fabrication are done at a constant rate (constrained
by the feed material availability).  The enrichment of the \gls{UOX} is processed
on demand.  Buffer storages (not shown in Figure \ref{fig:cycle}) are present
between all facilities with constant processing rates.

@@

It must be noted that for this work, the fuel enrichment was determined by a
fuel fabrication model\cite{Leniau2015125} which
calculates the fissile fraction required in the fuel to reach a target
burnup at the end of irradiation.  Therefore, the uncertainty of the fuel
enrichment is not applied directly to the effective enrichment. It is instead applied to the
burnup, which leads to a fuel enrichment uncertainty close to
$10\%$.

@@

The main interests of nuclear archaeology involve highly enriched uranium production and
fissile inventories, so this study mainly focuses on the uncertainties of
the total quantity of natural uranium used and the total unused fissile
inventory (unused reprocessed plutonium).  
Generated power is also briefly being investigated to ensure the overall
uncertainty is not contaminated with a missed fuel loading, which could
be interesting but is not the focus of this work.

For this study, the 199 simulations were performed six times, and a reference calculation was also
computed. With the five parameters in Table
\ref{tab:package_uncertainty} (separation efficiency, residence time, \gls{UOX} fuel
enrichment, reactor cycle length, reactor thermal power) being tracked, the six sets are divided as follows: five sets where a single parameter is varied and 
the four others remain fixed, and one with all five parameters being varied. 
The reference calculation is a single
simulation with all five parameters fixed at the mean value.

In the following, each total uncertainty is reported at $\pm1\sigma$ around the
mean values of the corresponding uncertainty set at each time step.  The 
time-dependant relative uncertainties are computed as the standard deviation of a
single set over its mean value at the corresponding time.

@@

By construction, the very low uncertainty (Figure \ref{fig:power_full}) on the
generated power is expected.

\begin{figure}[h!] % replace 't' with 'b' to force it to be on the bottom
    \centering
    \includegraphics[scale=0.36]{power_full}
    \caption{Electric power generated as a function of time.  The black line
        represents the reference calculation and the blue zone
        represents the 1 $\sigma$ uncertainty distribution.}\label{fig:power_full}
\end{figure}

Indeed the transition scenario has been designed to allow all the reactors to
receive the fuel as needed allowing them to be producing power all the time.
This ensures that the uncertainty measured any other parameters is not
influenced by the reactors miss-loadings.  As \glspl{FBR} fuel is built from
reprocessed used fuels, the availability of those used fuels may impact the
capability to build the required MOX-FBR fuel.  Having reactor miss-loads, may
impact more miss-loading later during a simulation and lead to a snow-ball
effect.  Having a small uncertainty on the power generation ensures large miss
load effects are not to be considered here.

@@

As expected, cooling time, separation efficiency and thermal power don't
affect the natural uranium consumption.  The uncertainty on the natural uranium
consumption is dominated by the \glspl{LWR} cycle length: shorter the cycle
length is, the more fuel will be loaded.

@@

The decrease of the cycle time uncertainty contribution, from year 30 to year
90, can be understand as an averaging effect.  As the uncertainty are considered
as systematic per deployed facility, each reactor will have a different cycle
length value, so a different number of loaded batches per year.With the increase
of deployed reactors, the number of batches of \glspl{LWR} fuel loaded converges
to the references one, reducing its impact on the uncertainty on the total
uranium consumption.

The fuel enrichment contribution follow a close to constant uncertainty
contribution of $2\%$, so as the total relative uncertainty decrease, its
relative contribution increases over time.

\subsection{Fissile Inventory}
The fissile inventory corresponds to the amount on separated fissile waiting to
be blend with natural uranium in order to produce the \glspl{FBR} MOX fuel.

As observed on Figure \ref{fig:pu_full}, the fissile inventory starts to grow on
year 40, with the deployment of the first reprocessing facility.  It grows
during the first period of the \glspl{LWR} to \glspl{FBR}-A transition until
year 90, then decrease slightly as the last \glspl{LWR} are decommissioned.
When he steady state is reached (full \glspl{FBR} fleet), the plutonium
inventory starts to increase, as the breading ratio of the \glspl{FBR} is higher
than the deployment needs.

@@

Regarding the uncertainty, beside the low inventory artefact observable around
year 40, the relative uncertainty starts at about $5\%$, and grows until year
125 to decrease up to $10\%$ at the end of the simulation time (180y), it seems
it would have reached an equilibrium between $5$ and $10\%$ with longer
simulations...
Between year 40 and 70, the main contributor to the fissile inventory
uncertainty is the cycle length, as for the uranium needs, lower cycle length
implies more fuel loads.  Never the less, around year 65 (see Figure
\ref{fig:used_fuel}), the stock of used fuel available for reprocessing starts
to lack, and the availability of the used fuel starts to dictate the evolution
of fissile in storage.  The uncertainty weight is then slowly transferred from
the cycle length to the fuel cooling time.

\begin{figure}[ht] % replace 't' with 'b' to force it to be on the bottom
    \centering
    \includegraphics[scale=0.18]{used_fuel}
    \caption{Used fuel inventory, corresponding to the amount of used
    \glspl{LWR} and \glspl{FBR} fuel, cooled and waiting to be reprocessed.}
    \label{fig:used_fuel}
\end{figure}
Between year 90 and 120, the \glspl{FBR} deployment speed increases, it has to
compensate the non replacement of the decommissioned \glspl{LWR} as well as
follow the power demand.  During this period the needs in plutonium are higher
than its productions, explaining the observed decrease of the total amount, and
the increase associated relative uncertainty.

As the transition ends, the plutonium breed from the \glspl{FBR} starts to be
sufficient to sustain the new \glspl{FBR} deployment to follow the power demand.

>>>>

\usepackage[acronym]{glossaries}
\newacronym{UOX}{UOX}{uranium oxide fuel}
\newacronym{MOX}{MOX}{mixed oxide fuel}
\newacronym{LWR}{LWR}{light water reactor}
\newacronym{SFR}{SFR}{fast breeder reactor}

@@

Fuel cycle simulations, as with any simulation process, do not produce results
without errors.  Those errors have different sources: data uncertainties (from
%% I'm not a fan of the use of the word ``errors'' here.  We should be careful
%% since (in English) this can meam ``mistakes''.  Uncertainty is a better
%% word, and then we'll have to distinguish between input uncertainties and
%% output uncertainties.
the simulator's reliance on previously estimated/measured data), modeling
uncertainties (from simplifications made by the simulator) and the
``operational'' uncertainty (from the expected behavior of the facilities
being simulated).
%% is this operational uncertainty really just a different kind of data?

Simulation errors could be important in the context of treaty verification. 
Some scenarios require estimating historical fissile
material production based on records of facility operation and material
transfer.  If the measurements of facility operation and material transfer are
subject to uncertainty, the total production quantities will then also be
uncertain.  Any uncertainty in fissile production quantities poses a material
accounting challenge, possibly
creating an opportunity for undetected diversion.  This study seeks
to understand which measures of facility operation have the most impact on the
uncertainty of fissile material production across a complete fuel cycle.
To accomplish this, a Total Monte Carlo methodology will be applied to the Cyclus
fuel cycle simulator \cite{cyclus} to estimate the effect of individual facility uncertainties on the output metrics.
%% should there be a reference for Total Monte Carlo?

@@

To estimate the uncertainty of fuel cycle output metrics, a Total Monte
Carlo approach has been applied.  In each of 199 independent simulations,
some of the parameters that define facility behavior are randomly selected
from Normal distributions with standard deviations that are $10\%$ of the mean
value.  This artificially large standard deviation is selected for the purpose
of demonstrating the methodology.  The mean and standard deviation of some
output metrics are calculated from the simulation output, as a function of time.

To pursue this work, both the Cycamore\footnote{Low
fidelity facilities package for Cyclus}\cite{cycamore} and the
CyCLASS\footnote{Reactor and fuel fabrication facility based on CLASS\cite{CLASS} models}\cite{mouginot_2018, cyclass} packages were
updated to evaluate the uncertainty of several
operational parameters in their associated facilities.  Table
\ref{tab:package_uncertainty} summarizes all uncertainty-tracking modifications implemented in the
facilities of Cycamore and CyCLASS.

\begin{table}[htb]
\centering
  \caption{Summary of facility modifications according to their Cyclus archetype package.}
\begin{tabular}{cll}
\toprule

Package   & Facility   & Parameters                \\
\midrule
Cycamore & Separation & Separation efficiency     \\ %% add hline in 2nd two columns
         & Storage    & Residence Time            \\
\midrule
CyCLASS  & Reactor    & Cycle Length              \\
         &            & Power                     \\
         &            & Fuel Enrichment (PWR-UOX) \\

@@

The uncertainty of the different parameters in this study
are systematic for each new deployed facility: each time a new facility is
deployed (such as a reactor or a fuel fabrication facility) a new set of
parameters are sampled, and those parameters are used throughout the life
of that facility.

\subsection{The Fuel Cycle Scenario}
For demonstration purposes, a simple commercial fuel cycle transition is used,
inspired from the EG23 fuel cycle
of the Nuclear Fuel Cycle Evaluation and Screening Report\cite{ES}.
Pictured in Figure \ref{fig:cycle}, this is a
transition from a \gls{LWR} fleet loaded with \gls{UOX} fuel to a
\gls{SFR} fleet loaded \gls{MOX} fuel, considering a $1\%/$y growth
of the nuclear generated power (Figure \ref{fig:power}).
\begin{figure}[ht] % replace 't' with 'b' to force it to be on the bottom
  \centering
  \includegraphics[scale=0.31]{cycle.png}
  \caption{Simplified representation of the material flow between the different facilities,
  with each operational parameter associated with an
  uncertainty labeled in green.}\label{fig:cycle}
\end{figure}
%% Update this to match the one in the CVT poster?

As illustrated in Figure \ref{fig:power}, the transition starts with a fleet
composed of only \glspl{LWR} loaded with enriched \gls{UOX} fuel.  The actual
transition starts around year 35, with the deployment of \glspl{SFR}. The \glspl{SFR} are loaded
with \gls{MOX}, which consists of plutonium blended with natural
uranium.

\begin{figure}[ht] % replace 't' with 'b' to force it to be on the bottom
    \centering
    \includegraphics[scale=0.18]{power.png}
    \caption{Electric power generated over time, with the full production,
        the \gls{LWR} contribution, and the \gls{SFR} contribution labeled in blue, orange, and green, respectively.}
    \label{fig:power}
\end{figure}

@@

Both fuel reprocessing and fabrication are done at a constant rate (constrained
by the feed material availability).  The enrichment of the \gls{UOX} is processed
%% is it really a constant rate then?  Maybe we need a better word to describe this
on demand.  Buffer storage facilities (not shown in Figure \ref{fig:cycle}) are present
%% You can't pluralize ``storage'' in English - it's considered uncountable
between all facilities with constant processing rates.

@@

It must be noted that for this work, the fuel enrichment was determined by a
fuel fabrication model\cite{Leniau2015125} which
calculates the fissile fraction required in the fuel to reach a given target
burnup at the end of irradiation.  Therefore, the uncertainty of the fuel
enrichment is not applied directly to the effective enrichment. It is instead applied to the
burnup, which leads to a fuel enrichment uncertainty close to
$10\%$.
%% Did you have to experiment to find the burnup uncertainty that results in a 10% enrichment uncertainty?

For each of the parameters in Table \ref{tab:package_uncertainty}, a set of
199 simulations was performed with only that parameter being randomly varied.
An additional set of 199 simulations was performed in which all five
parameters are randomly varied.  A reference simulation was performed
with all five parameters fixed at the mean value.

@@

Because the main interests of nuclear archaeology involve highly enriched uranium production and
fissile inventories, this study mainly focuses on the uncertainties of
the total quantity of natural uranium used and the total unused fissile
inventory (unused reprocessed plutonium).
%% unused implies it's never used, but I don't think that's right.  You just mean the amount
%% that is being stored in a separated state any point in time, right?
Generated power is also briefly investigated to ensure the overall
uncertainty is not contaminated with a missed fuel loading, which could
be interesting but is not the focus of this work.
%% what do you mean by contaminated???  This probably deserves more explanation

At each time step, the $\pm1\sigma$ uncertainty due to the full set of varied
parameters is reported.  In addition, the relative contribution to the total
uncertainty from each of the parameters is calculated as the ratio of the
$1\sigma$ uncertainty due to that parameter over the total uncertainty.

@@

\begin{figure}[t] % replace 't' with 'b' to force it to be on the bottom
    \centering
    \includegraphics[scale=0.36]{power_full}
    \caption{Electric power generated as a function of time.  The black line
        represents the reference calculation and the blue zone
        represents the 1 $\sigma$ uncertainty distribution.}\label{fig:power_full}
\end{figure}

By design, the very low uncertainty (Figure \ref{fig:power_full}) on the
generated power is expected.
Indeed the transition scenario has been designed to allow all the reactors to
receive the fuel as needed allowing them to be producing power all the time.
This ensures that the measured uncertainty is not affected by a facility
disruption that has an outsized impact on material flows.
As \gls{SFR} fuel is built from
reprocessed used fuel, the availability of that used fuel may impact the
capability to build the required MOX-SFR fuel.  Facility disruptions may
cause material shortages that can lead to snow-ball effect of more disruptions
later in the simulation.  While such disruptions are interesting in some analyses,
they are outside the scope of this study.

@@

As expected, cooling time, separation efficiency and thermal power do not
affect the natural uranium consumption.  The uncertainty on the natural uranium
consumption is dominated by the \gls{LWR} cycle length: shorter the cycle
length is, the more fuel will be loaded.

@@

The decrease of the cycle time uncertainty contribution, from year 30 to year
90, can be understood as an averaging effect.  As the uncertainty are defined once
per deployed facility, each reactor will have a different cycle
length value, so a different number of batches loaded per year.  With the increase
of deployed reactors, the number of batches of \gls{LWR} fuel loaded converges
to the references one, reducing its impact on the uncertainty on the total
uranium consumption.
%% Why does the relative uncertainty due to cyle time not approach 0?

The fuel enrichment contribution follow a close to constant uncertainty
contribution of $2\%$, so as the total relative uncertainty decrease, it
has a growing shared of the total uncertainty over time.

\subsection{Fissile Inventory}
The fissile inventory corresponds to the amount of separated fissile material waiting to
be blended with natural uranium in order to produce the \gls{SFR} MOX fuel.

As observed on Figure \ref{fig:pu_full}, the fissile inventory starts to grow in
year 40, with the deployment of the first reprocessing facility.  It grows
during the first period of the \glspl{LWR} to \gls{SFR}-A transition until
year 90, then decrease slightly as the last \glspl{LWR} are decommissioned.
When he steady state is reached (full \gls{SFR} fleet), the plutonium
inventory starts to increase, as the breading ratio of the \glspl{SFR} is higher
than the deployment needs.

@@

Regarding the uncertainty, disregarding the artefact of a very low mean inventory
observed at the onset of fissile material production, the relative uncertainty
starts at about $5\%$, and grows until year
125 to decrease up to $10\%$ at the end of the simulation time (180y), it seems
%% I don't see what you mean by decrease up to 10%?
it would have reached an equilibrium between $5$ and $10\%$ with longer
simulations...
Between year 40 and 70, the main contributor to the fissile inventory
uncertainty is the cycle length.  As with the uranium needs, lower cycle length
implies more fuel loads.  Never the less, around year 65 (see Figure
%% Aren't you varying the cycle length equally to be smaller and larger, so why
%% focus on ``lower cycle lengths''?
\ref{fig:used_fuel}), the stock of used fuel available for reprocessing becomes
small enough to start limiting producton of fissile material.
The uncertainty weight is then slowly transferred from
the cycle length to the fuel cooling time.

\begin{figure}[ht] % replace 't' with 'b' to force it to be on the bottom
    \centering
    \includegraphics[scale=0.18]{used_fuel}
    \caption{Used fuel inventory, corresponding to the amount of used
    \glspl{LWR} and \glspl{SFR} fuel, cooled and waiting to be reprocessed.}
    \label{fig:used_fuel}
\end{figure}
Between year 90 and 120, the \gls{SFR} deployment speed increases, it has to
compensate the non replacement of the decommissioned \glspl{LWR} as well as
follow the power demand.  During this period the needs in plutonium are higher
than its productions, explaining the observed decrease of the total amount, and
the increase associated relative uncertainty.

As the transition ends, the plutonium breed from the \glspl{SFR} starts to be
sufficient to sustain the new \glspl{SFR} deployment to follow the power demand.
